The very dependent recursive structure of iterated parametricity in indexed form

### 2025-12-04 paper/paper.tex
paper: drop some univ subscripts
Many fixes in paper

\begin{document}
\title{On the specific recursive structure of iterated parametricity}
\begin{authgrp}
  \author{Hugo Herbelin}
  \affiliation{Université Paris Cité, Inria, CNRS, IRIF, Paris
    \email{Hugo.Herbelin@inria.fr}}
  \author{Ramkumar Ramachandra}
  \affiliation{Unaffiliated \email{r@artagnon.com}}
\end{authgrp}
\begin{abstract}
  In a previous paper, we gave equations describing the construction of semi-simplicial and semi-cubical sets via iterated parametricity. The construction was formalised in the Rocq (formerly, Coq) proof assistant building $n$-truncated sets using an induction on a large structure embedding equational reasoning.

@@

\begin{align*}
  \framep[n][p]   & : \U[m]                                                                                                     & \mbox{for all $p\leq n$} \\
  \painting[n][p] & : \framep[n][p] \rightarrow \U[m]                                                                           & \mbox{for all $p\leq n$} \\
  \restrf[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]                               & \mbox{for all $p\leq n$} \\
  \restrp[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                                           \\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                  & \mbox{for all $p\leq n$} \\
  \cohf[n][p]     & : \Pi qr\epsilon\omega.\, \restrf[n][p][\epsilon][q] \circ \restrf[n+1][p][\omega][r] =                                                \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][\epsilon][q+1]                                       & \mbox{for all $p\leq n$} \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,                                                                                        \\
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d) & \mbox{for all $p\leq n$} \\
\end{align*}

@@

The definition of $\framep[n+1][p]$ also depends, via $\layer[n][p-1]$, on $\painting[n][p-1]$. Both via $\painting[n][p-1]$ and $\restrf[n][p-1]$, it depends also on $\framep[n][p-1]$ by typing dependency. This requires to assume $\framep[n][p-1]$ and $\painting[n][p-1]$ known in advance.
Moreover, the definition of $\framep[n+1][p]$ depends on $\framep[n+1][p-1]$, so all of $\framep[n][k]$, $\painting[n][k]$ and $\restrf[n][k]$ for $k< p$ need to be known, which we respectively abbreviate as $\framep[n][0,...,p-1]$, $\painting[n][0,...,p-1]$ and $\restrf[n][p-1]$.
Schematically, using dotted arrows for dependencies in the type and plain arrows for dependencies in the definition, this gives the following dependencies for $p\leq n$ fixed:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    \framep[n+1][0,...,p] \arrow[d] \arrow[ddr]\\
    \painting[n][0,...,p-1] \arrow[d, dotted] \\
    \framep[n][0,...,p-1] &
    \restrf[n][0,...,p-1] \arrow[l, dotted] \arrow[uul, dotted, "\mbox{$[0,...,p-1]$}"', near start, shift right=2mm, hook']\\
  \end{tikzcd}
\end{center}

@@

Assuming $\framep[n][0,...,p]$ only requires to define the type of frames, which are just finite products of $\U$:

@@

The types $\paintingtype[n][0,...,p]$ of $\painting[n][0,...,p]$
are type families over frames, so $\framep[n][0,...,p]$ needs to be assumed
beforehand:

@@

Assuming $\framep[n][0,...,p-1]$ and $\painting[n][0,...,p-1]$, we can
mutually define the types $\restrftype[n][0,...,p-1]$ of frame
restrictions together with $\framep[n+1][0,...p]$ as a function from
$\restrf[n][0,...,p-1]$, introducing also abbreviations for the types
and definitions of layers, as well as for the combination of sequences
of $\framep[n][0,..,p-1]$, $\painting[n][0,...,p-1]$ and
$\restrf[n][0,...,p-1]$ in one block, which we call
$\depstype[n][0,...p-1]$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \depstype[n][0,..,p-1]          &        & \defeq            &
    \left(\begin{array}{lll}
              \framep[n][0,...,p-1]   & : & \frametype[n][0,...,p-1]    \\
              \painting[n][0,...,p-1] & : & \paintingtype[n][0,...,p-1]
              (\framep[n][0,...,p-1])                                   \\
              \restrf[n][0,...,p-1]   & : & \restrftype[n][0,...,p-1]
              \left(\begin{array}{l}
                  \framep[n][0,...,p-1]   \\
                  \painting[n][0,...,p-1] \\
                \end{array}\right)                             \\
            \end{array}\right)        \\
    \framep[n+1][0,...,(p+1)-1]     &
    \multicolumn{3}{l}{
    (\deps[n][0,...,p-1]:\depstype[n][0,...,p-1]) :\frametype[n+1][0,...,p]}         \\
    \framep[n+1][0,...,(0+1)-1]     &
    (\unitpoint)                    & \defeq & (\unitpoint,unit)                     \\
    \framep[n+1][0,...,((p+1)+1)-1] &
    \deps[n][0,...,p]               & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][0,...,p]
              (\deps[n][0,...,p-1]), \\
              \Sigma d:\framep[n+1][0,...,p]
              (\deps[n][0,...,p-1])  \\
              \quad \layer[n][p]
              (\deps[n][0,...,p])
              (d)                    \\
            \end{array}\right)                                              \\
    \restrftype[n][0,...,p-1]       &
    \multicolumn{3}{l}{
    \left(\begin{array}{l}
              \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
              \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
            \end{array}\right) : \U} \\
    \restrftype[n][0,...,0-1]       &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)          & \defeq &
    \unittype                                                                        \\
    \restrftype[n][0,...,(p+1)-1]   &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
            \end{array}\right)        & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][0,...,p-1]
      :\restrftype[n][0,...,p-1]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
              \end{array}\right).                     \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
                \restrf[n][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}               \\
    \layertype[n][p]                &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
    : \U}                                                                            \\
    \layertype[n][p]                &
    \deps[n][0,...,p]
                                    & \defeq & \framep[n+1][p]
    (\deps[n][0,...,p-1])
    \rightarrow \U                                                                   \\
    \layer[n][p]                    &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
      : \layertype[n][p]
      (\deps[n][0,...,p])}
    \\
    \layer[n][p]                    &
    \deps[n][0,...,p]
                                    & \defeq &
    d \mapsto \Pi\omega.\,\painting[n][p](\restrf[n][p](d))                          \\
  \end{array}
\end{equation*}

Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the projections of a $\deps$. Note that by abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Defining $\painting[n+1][p]$, in addition to the type dependency in $\framep[n+1][p]$, requires the definition of $\painting[n][p-1]$, thus also $\framep[n][p-1]$ by typing dependency, as well as $\restrf[n][p-1]$. Since $\painting[n+1][p]$ also depends on $\painting[n+1][p+1]$ when $p<n+1$, these dependencies are actually between $\painting[n+1][p,...,n+1]$, $\painting[n][p,...,n]$, $\framep[n+1][p,...,n+1]$, $\framep[n][p,...,n]$ and $\restrf[n][p,...,n]$. Also, for the case $p=n$, this requires the assumption of an inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally, $\framep[n+1][p,...,n+1]$, $\framep[n+1][p,...,n]$ and $\restrf[n][p,...,n]$ respectively depend on the earlier values $\framep[n+1][0,...,p-1]$, $\framep[n][0,...,p-1]$ and $\restrf[n][0,...,p-1]$.  Below, we use dashed underline and dashed arrows for dependencies in subconstructions, here in the construction of $\framep$ at level $n$ and $\restrftype$ at level $n$:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    & & \painting[n+1][p,...,n+1] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
    \udensdash{\framep[n+1][0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
    \udensdash{\framep[n+1][p,...,n+1]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
    \painting[n][0,...,p-1] \arrow[d, dotted] & &
    \painting[n][p,...,n-1] \arrow[ll, dotted] \arrow[d, dotted] \\
    \framep[n][0,...,p-1] &
    \udensdash{\restrf[n][0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
    \framep[n][p,...,n] \arrow[ll, dotted, bend left=15] &
    \udensdash{\restrf[n+1][p,...,n]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
  \end{tikzcd}
\end{center}

Let us introduce another abbreviation to combine a full $\deps[n][0,...,n-1]:\depstype[n][0,...,n-1]$ with a family $E:\framep[n+1][n](\deps[n][0,...,n-1])\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
            \deps[n][0,...,n-1] & : & \depstype[n][0,...,n-1]                             \\
            E                   & : & \framep[n+1][n](\deps[n][0,...,n-1]) \rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

We can now formally define $\painting[n+1][p]$ by recursion on $n+1-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][p]      &
    (\fulldeps[n]:\fulldepstype[n])                           \\
    \painting[n+1][n]      &
    \fulldeps[n]           & \defeq & \fulldeps.E             \\
    \painting[n+1][p](p<n) &
    \fulldeps[n]           & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][0,...,p]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                 \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

From which we can deduce $\painting[n+1][0,...,p]$ by a second recursion, this time on $p$:

@@

At this stage, we are able to define $\framep[n+1][0,...,n]$, $\painting[n+1][0,...,n]$ and $\restrftype[n][0,...,n-1]$ from $\framep[n][0,...,n-1]$, $\painting[n][0,...,n-1]$ and $\restrf[n][0,...,n-1]$, that is from $\deps[n][0,...,n-1]$. To go further, we need to define the sequence $\restrptype[n][0,...,p]$ of types of painting restrictions:

@@

Thanks to $\restrptype[n][0,...,p-1]$, we can mutually define $\restrf[n+1][0,...,p]$ and $\cohftype[n][0,...,p-1]$ from $\framep[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$, $\restrptype[n][0,...,p-1]$ and some $E:\framep[n+1][n](\deps[n][0,...,n-1]) \rightarrow \U$, abbreviating these dependencies and the dependency in a $\cohf[n][0,...,p-1]$ as $\depscohttype[n][0,...,p-1]$, and using also abbreviations in passing for $\restrl$ and $\restrltype$. It is also convenient to build $\deps[n+1][0,...,p]$ from $\depscohs[n][0,...,p-1]$, so that we can now build $\framep[n+2][p]$ one step further. Thus, we also include this abbreviation in the definition:

@@

    \restrf[n+1][0,...,(p+1)-1]     &

@@

    \restrf[n+1][0,...,(0+1)-1]     &

@@

    \restrf[n+1][0,...,((p+1)+1)-1] &

@@

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n+2][n+1](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{lcl}
            \depscohs[n][0,...,n-1] & : & \depscohstype[n][0,...,n-1]                                                \\
            E                       & : & \framep[n+2][n+1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \fulldepstype[n+1] \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs[n][0,...,n-1]) \\
            \fulldepscohs.\depscohs.\fulldeps.E                     \\
          \end{array}\right)                             \\
\end{align*}

Also, when $n$ is 0, we write $(E',E)$ to denote an element $(E', E)$ of $\fulldepscohstype[0]$.

@@

Armed with $\restrf$, $\cohftype$ and these abbreviations, it is now possible to define $\restrp[n][p]$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][p]                         &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n+2][p]
        (\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])). \\
        \painting[n+2][p]
        (\overline{\fulldeps}(\fulldepscohs[n]))(d) \rightarrow    \\
        \painting[n+1][p](\fulldepscohs.\depscohs.\fulldeps[n])
        (\restrf[n+1][p](\fulldepscohs.\depscohs[n][0,...,p-1])(d))\end{array}\right)                                                        \\
    \restrp[n+1][n]                         &
    (\fulldepscohs[n])                      & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n+1][p](p<n)                    &
    (\fulldepscohs[n])                      & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n][0,...,p])(d)(l), \\
              \restrp[n+1][p+1][q][\epsilon](\fulldepscohs[n])(d,l)(c)               \\
            \end{array}\right)                                             \\
  \end{array}
\end{equation*}

From which we can deduce $\restrp[n][0,...,p]$ again by a second recursion, this time on $p$:

@@

At this stage, we are then able to define $\restrf[n+1][0,...,p]$, $\restrp[n+1][0,...,p]$ and $\cohftype[n][0,...,p]$
from $\depscohs[n][0,...,p-1]$ at the previous level. To go further, we need to define the
sequence $\cohptype[n][0,...,p-1]$ of types of painting coherence laws:

@@

Thanks to $\cohptype[n][0,...,p-1]$, we can mutually define $\cohf[n+1][0,...,p]$ and $\cohttype[n][0,...,p-1]$ from $\fulldepscohs[n][0,...,p-1]$, abbreviating this and the dependency in a $\coht[n][0,...,p-1]$ as $\depscohttype[n][0,...,p-1]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][0,...,p]$ from $\depscoht[n][0,...,p-1]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

@@

    \cohf[n][0,...,((p+1)+1)-1]                                      &
    (\fulldepscoht[n])                                               & \defeq                                            &

>>>>

\begin{document}
\title{The very dependent recursive structure of iterated parametricity in indexed form}
\begin{authgrp}
  \author{Hugo Herbelin}
  \affiliation{Université Paris Cité, Inria, CNRS, IRIF, Paris
    \email{Hugo.Herbelin@inria.fr}}
  \author{Ramkumar Ramachandra}
  \affiliation{Unaffiliated \email{r@artagnon.com}}
\end{authgrp}
\begin{abstract}
  In a previous paper, we gave equations describing the construction of semi-simplicial and semi-cubical sets via iterated parametricity. The construction was formalised in the Rocq (formerly, Coq) proof assistant building $n$-truncated sets using an induction on a large structure embedding equational reasoning.

@@

\begin{align*}
  \framep[n][p]   & : \U                                                                                                        & \mbox{for all $p\leq n$} \\
  \painting[n][p] & : \framep[n][p] \rightarrow \U                                                                              & \mbox{for all $p\leq n$} \\
  \restrf[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]                               & \mbox{for all $p\leq n$} \\
  \restrp[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                                           \\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                  & \mbox{for all $p\leq n$} \\
  \cohf[n][p]     & : \Pi qr\epsilon\omega.\, \restrf[n][p][\epsilon][q] \circ \restrf[n+1][p][\omega][r] =                                                \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][\epsilon][q+1]                                       & \mbox{for all $p\leq n$} \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,                                                                                        \\
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d) & \mbox{for all $p\leq n$} \\
\end{align*}

@@

The definition of $\framep[n+1][p]$ also depends, via $\layer[n][p-1]$, on $\painting[n][p-1]$. Both via $\painting[n][p-1]$ and $\restrf[n][p-1]$, it depends also on $\framep[n][p-1]$ by typing dependency. This requires to assume $\framep[n][p-1]$ and $\painting[n][p-1]$ known in advance.
Moreover, the definition of $\framep[n+1][p]$ depends on $\framep[n+1][p-1]$, so all of $\framep[n][k]$, $\painting[n][k]$ and $\restrf[n][k]$ for $k< p$ need to be known, which we respectively abbreviate as $\framep[n][0,...,p-1]$, $\painting[n][0,...,p-1]$ and $\restrf[n][p-1]$.
Schematically, using dotted arrows for dependencies in the type and plain arrows for dependencies in the definition, this gives the following dependencies for $p\leq n+1$ fixed:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    \framep[n+1][0,...,p] \arrow[d] \arrow[ddr]\\
    \painting[n][0,...,p-1] \arrow[d, dotted] \\
    \framep[n][0,...,p-1] &
    \restrf[n][0,...,p-1] \arrow[l, dotted] \arrow[uul, dotted, "\mbox{$[0,...,p-1]$}"', near start, shift right=2mm, hook']\\
  \end{tikzcd}
\end{center}

@@

Assuming $\framep[n][0,...,p]$ for $p \leq n$ only requires to define the type of frames, which are just finite products of $\U$:

@@

The types $\paintingtype[n][0,...,p]$ of $\painting[n][0,...,p]$ for $p \leq n$
are type families over frames, so $\framep[n][0,...,p]$ needs to be assumed
beforehand:

@@

Assuming $\framep[n][0,...,p-1]$ and $\painting[n][0,...,p-1]$ for $p \leq n+1$, we can
mutually define the types $\restrftype[n][0,...,p-1]$ of frame
restrictions together with $\framep[n+1][0,...p]$ as a function from
$\restrf[n][0,...,p-1]$, introducing also abbreviations for the types
and definitions of layers, as well as for the combination of sequences
of $\framep[n][0,..,p-1]$, $\painting[n][0,...,p-1]$ and
$\restrf[n][0,...,p-1]$ in one block, which we call
$\depstype[n][0,...p-1]$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \depstype[n][0,..,p-1]          &        & \defeq            &
    \left(\begin{array}{lll}
              \framep[n][0,...,p-1]   & : & \frametype[n][0,...,p-1]    \\
              \painting[n][0,...,p-1] & : & \paintingtype[n][0,...,p-1]
              (\framep[n][0,...,p-1])                                   \\
              \restrf[n][0,...,p-1]   & : & \restrftype[n][0,...,p-1]
              \left(\begin{array}{l}
                  \framep[n][0,...,p-1]   \\
                  \painting[n][0,...,p-1] \\
                \end{array}\right)                             \\
            \end{array}\right)        \\
    \framep[n+1][0,...,p]     &
    \multicolumn{3}{l}{
    (\deps[n][0,...,p-1]:\depstype[n][0,...,p-1]) :\frametype[n+1][0,...,p]}         \\
    \framep[n+1][0,...,0]     &
    \unitpoint                    & \defeq & (\unitpoint,unit)                     \\
    \framep[n+1][0,...,p+1] &
    \deps[n][0,...,p]               & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][0,...,p]
              (\deps[n][0,...,p-1]), \\
              \Sigma d:\framep[n+1][p]
              (\deps[n][0,...,p-1]).
              \, \layer[n][p]
              (\deps[n][0,...,p])
              (d)                    \\
            \end{array}\right)                                              \\
    \restrftype[n][0,...,p-1]       &
    \multicolumn{3}{l}{
    \left(\begin{array}{l}
              \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
              \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
            \end{array}\right) : \U} \\
    \restrftype[n][0,...,0-1]       &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)          & \defeq &
    \unittype                                                                        \\
    \restrftype[n][0,...,(p+1)-1]   &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
            \end{array}\right)        & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][0,...,p-1]
      :\restrftype[n][0,...,p-1]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
              \end{array}\right).                     \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
                \restrf[n][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}               \\
    \layertype[n][p]                &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
    : \U}                                                                            \\
    \layertype[n][p]                &
    \deps[n][0,...,p]
                                    & \defeq & \framep[n+1][p]
    (\deps[n][0,...,p-1])
    \rightarrow \U                                                                   \\
    \layer[n][p]                    &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
      : \layertype[n][p]
      (\deps[n][0,...,p])}
    \\
    \layer[n][p]                    &
    \deps[n][0,...,p]
                                    & \defeq &
    d \mapsto \Pi\omega.\,\painting[n][p](\deps.\restrf[n][p][0][\omega](d))                          \\
  \end{array}
\end{equation*}

Note that we write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the
projections of a $\deps$. Also, by abuse of notation, we write $\unitpoint$ for
the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Defining $\painting[n+1][p]$ for $p \leq n+1$, in addition to the type dependency in $\framep[n+1][p]$, requires the definition of $\painting[n][p-1]$, thus also $\framep[n][p-1]$ by typing dependency, as well as $\restrf[n][p-1]$. Since $\painting[n+1][p]$ also depends on $\painting[n+1][p+1]$ when $p<n+1$, these dependencies are actually between $\painting[n+1][p,...,n+1]$, $\painting[n][p,...,n]$, $\framep[n+1][p,...,n+1]$, $\framep[n][p,...,n]$ and $\restrf[n][p,...,n]$. Also, for the case $p=n$, this requires the assumption of an inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally, $\framep[n+1][p,...,n+1]$, $\framep[n+1][p,...,n]$ and $\restrf[n][p,...,n]$ respectively depend on the earlier values $\framep[n+1][0,...,p-1]$, $\framep[n][0,...,p-1]$ and $\restrf[n][0,...,p-1]$.  Below, we use dashed underline and dashed arrows for dependencies in subconstructions, here in the construction of $\framep$ at level $n$ and $\restrftype$ at level $n$:
\begin{center}
  \fontsize{7.2}{9}\selectfont
  \begin{tikzcd}
    & & \painting[n+1][p,...,n+1] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
    \udensdash{\framep[n+1][0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
    \udensdash{\framep[n+1][p,...,n+1]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
    \painting[n][0,...,p-1] \arrow[d, dotted] & &
    \painting[n][p,...,n-1] \arrow[ll, dotted] \arrow[d, dotted] \\
    \framep[n][0,...,p-1] &
    \udensdash{\restrf[n][0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
    \framep[n][p,...,n] \arrow[ll, dotted, bend left=15] &
    \udensdash{\restrf[n+1][p,...,n]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
  \end{tikzcd}
\end{center}

Let us introduce another abbreviation to combine a full $\deps[n][0,...,n]:\depstype[n][0,...,n]$ with a family $E:\framep[n+1][n+1](\deps[n][0,...,n])\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
            \deps[n][0,...,n] & : & \depstype[n][0,...,n]                             \\
            E                   & : & \framep[n+1][n+1](\deps[n][0,...,n]) \rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

We can now formally define $\painting[n+1][p]$ by recursion on $n+1-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][p]      &
    (\fulldeps[n]:\fulldepstype[n])                           \\
    \painting[n+1][n+1]      &
    \fulldeps[n]           & \defeq & \fulldeps.E             \\
    \painting[n+1][p](p\leq n) &
    \fulldeps[n]           & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][0,...,p]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                 \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

From which we can deduce $\painting[n+1][0,...,p]$ for $p \leq n+1$ by a second recursion, this time on $p$:

@@

At this stage, we are able to define $\framep[n+1][0,...,n+1]$, $\painting[n+1][0,...,n+1]$ and $\restrftype[n][0,...,n]$ from $\framep[n][0,...,n]$, $\painting[n][0,...,n]$ and $\restrf[n][0,...,n]$, that is from $\deps[n][0,...,n]$. To go further, we need to define the sequence $\restrptype[n][0,...,p-1]$ of types of painting restrictions, for $p \leq n+1$:

@@

Thanks to $\restrptype[n][0,...,p-1]$ for $p \leq n+1$, we can mutually define $\restrf[n+1][0,...,p]$ and $\cohftype[n][0,...,p-1]$ from $\framep[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$, $\restrptype[n][0,...,p-1]$ and some $E:\framep[n+1][n+1](\deps[n][0,...,n]) \rightarrow \U$, abbreviating these dependencies and the dependency in a $\cohf[n][0,...,p-1]$ as $\depscohstype[n][0,...,p-1]$, and using also abbreviations in passing for $\restrl$ and $\restrltype$. It is also convenient to build $\deps[n+1][0,...,p]$ from $\depscohs[n][0,...,p-1]$, so that we can now build $\framep[n+2][p]$ one step further. Thus, we also include this abbreviation in the definition:

@@

    \restrf[n+1][0,...,p]     &

@@

    \restrf[n+1][0,...,0]     &

@@

    \restrf[n+1][0,...,p+1] &

@@

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n+2][n+2](\overline{\deps}(\depscohs[n][0,...,n])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{lcl}
            \depscohs[n][0,...,n-1] & : & \depscohstype[n][0,...,n]                                                \\
            E                       & : & \framep[n+2][n+2](\overline{\deps}(\depscohs[n][0,...,n]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \fulldepstype[n+1] \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs[n][0,...,n]) \\
            \fulldepscohs.E                     \\
          \end{array}\right)                             \\
\end{align*}

Also, when $n$ is 0, we write $(E,E')$ to denote an element $(E, E')$ of $\fulldepscohstype[0]$.

@@

Armed with $\restrf$, $\cohftype$ and these abbreviations, it is now possible to define $\restrp[n+1][p]$ for $p \leq n+1$ by recursion on $n-p$:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][p]                         &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n+2][p]
        (\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])). \\
        \painting[n+2][p]
        (\overline{\fulldeps}(\fulldepscohs[n]))(d) \rightarrow    \\
        \painting[n+1][p](\fulldepscohs.\depscohs.\fulldeps[n])
        (\restrf[n+1][p](\fulldepscohs.\depscohs[n][0,...,p-1])(d))\end{array}\right)                                                        \\
    \restrp[n+1][n+1]                         &
    (\fulldepscohs[n])                      & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n+1][p](p\leq n)                    &
    (\fulldepscohs[n])                      & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n][0,...,p])(d)(l), \\
              \restrp[n+1][p+1][q][\epsilon](\fulldepscohs[n])(d,l)(c)               \\
            \end{array}\right)                                             \\
  \end{array}
\end{equation*}

From which we can deduce $\restrp[n+1][0,...,p]$ for $p \leq n+1$ again by a second recursion, this time on $p$:

@@

At this stage, we are then able to define $\restrf[n+1][0,...,p]$, $\restrp[n+1][0,...,p]$ and $\cohftype[n][0,...,p-1]$ for $p \leq n+1$
from $\depscohs[n][0,...,p-1]$ at the previous level. To go further, we need to define the
sequence $\cohptype[n][0,...,p-1]$ of types of painting coherence laws:

@@

Thanks to $\cohptype[n][0,...,p-1]$, we can mutually define $\cohf[n+1][0,...,p]$ and $\cohttype[n][0,...,p-1]$ for $p \leq n+1$ from $\fulldepscohs[n][0,...,p-1]$, abbreviating this and the dependency in a $\coht[n][0,...,p-1]$ as $\depscohttype[n][0,...,p-1]$, and using also abbreviations in passing for $\cohl$ and $\cohltype$. It is also convenient to build $\depscohs[n+1][0,...,p]$ from $\depscoht[n][0,...,p-1]$, so that we can now build $\framep[n+3][p]$. Thus, we also include this abbreviation in the definition:

@@

    \cohf[n][0,...,p+1]                                      &
    (\depscoht[n][0,...,p-1])                                               & \defeq                                            &
